A Study of Common Bug Fix Patterns in Rust

### 2022-11-27 methodology.tex
double period elimination
fixes added

We propose a novel representation. Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn (the columns are constant across all Rust programs). However, to simulate order-sensitiveness, we add two more steps. First, we record the number of occurrences of items within the paths at each scope. The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension. Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs. Section~\ref{sec:weighting_scheme}, immediately below, illustrates this approach on an example..

@@

To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus. Specifically, we mined 20 of the most recent bug related commits of the top 20 most-starred Rust projects on GitHub, and ran them through our pipeline. We recorded the total number of occurrences of each item. Then, if $\# i$ is the number of occurences of $i$, we gave $i$ a weight of $1/\# i$. Because items that occur closer to the root also occur more often, per our observation, this gives higher weights to items that occur infrequently, i.e. closer to the leaves. Items that occur often and have lower weights are, we believe, less important in describing a change.

>>>>

We propose a novel representation. Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn (the columns are constant across all Rust programs). However, to simulate order-sensitiveness, we add two more steps. First, we record the number of occurrences of items within the paths at each scope. The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension. Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs. Section~\ref{sec:weighting_scheme}, immediately below, illustrates this approach on an example.

@@

To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus. Specifically, we mined 20 of the most recent bug related commits of the same projects described in Table~\ref{table:repos}, and ran them through our pipeline. We recorded the total number of occurrences of each item. Then, if $\# i$ is the number of occurences of $i$, we gave $i$ a weight of $1/\# i$. Because items that occur closer to the root also occur more often, per our observation, this gives higher weights to items that occur infrequently, i.e. closer to the leaves. Items that occur often and have lower weights are, we believe, less important in describing a change.

### 2022-11-27 results.tex
fixes added

We carried out a manual experiment for evaluating our code embedding. In the experiment, the human analyzer looks at the change's source code and finds the most important element that appears in the circle pack figure, recording its rank. A value of 1 indicates that the human analyzer concurs with the algorithm, while 2 means that the human-identified most important element was ranked 2 by the algorithm. After collecting all the ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$. Both authors acted as analyzer and carried out the experiment independently. To avoid inflating our final results, after comparing the ranks that the authors gave to each datapoint, we consider the larger (worse) rank in our final results. That is, if the first author gave rank 2 to datapoint $D$ and the second author gave rank 3 to that datapoint, in the final results we gave datapoint $D$ a rank of 3. The final results are shown in Figure~\ref{fig:rq1}.

@@

\addplot[
    color=green,
    mark=o,
    ]
    coordinates { 
    (1,64)
    (2,72)
    (3,84)
    (4,88)
    (5,92)
    };
    \addlegendentry{General}

\addplot[
    color=red,
    mark=square,
    ]
    coordinates { 
    (1,68)
    (2,84)
    (3,88)
    (4,92)
    (5,96)
    };
    \addlegendentry{Borrow-Checker}

\addplot[
    color=blue,
    mark=triangle,
    ]
    coordinates { 
    (1,66)
    (2,78)
    (3,86)
    (4,90)
    (5,94)
    };
    \addlegendentry{Total}

@@

Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 90\% of the time---we investigated the other 10\% as well. We found that our embedding tended to fail in change patterns where the change involves many elements (e.g. adding a whole new \verb+impl+ block and implementing new functions in it). It is worth mentioning that these change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.

>>>>

We carried out a manual experiment for evaluating our code embedding. In the experiment, the human analyzer looks at the change's source code and finds the most important element that appears in the circle pack figure, recording its rank. A value of 1 indicates that the human analyzer concurs with the algorithm, while 2 means that the human-identified most important element was ranked 2 by the algorithm. After collecting all the ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$. Both authors acted as analyzer and carried out the experiment independently. To avoid inflating our final results, after comparing the ranks that the authors gave to each datapoint, we consider the larger (worse) rank in our final results. That is, if the first author gave rank 2 to datapoint $D$ and the second author gave rank 3 to that datapoint, in the final results we gave datapoint $D$ a rank of 3. The authors' answers were the same for 45 of the 50 datapoints. The final results are shown in Figure~\ref{fig:rq1}.

@@

\addplot[
    color=green,
    mark=o,
    ]
    coordinates { 
    (1,72)
    (2,76)
    (3,88)
    (4,88)
    (5,92)
    };
    \addlegendentry{General}

\addplot[
    color=red,
    mark=square,
    ]
    coordinates { 
    (1,68)
    (2,84)
    (3,84)
    (4,88)
    (5,96)
    };
    \addlegendentry{Borrow-Checker}

\addplot[
    color=blue,
    mark=triangle,
    ]
    coordinates { 
    (1,70)
    (2,80)
    (3,86)
    (4,88)
    (5,94)
    };
    \addlegendentry{Total}

@@

Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 88\% of the time---we investigated the other 12\% as well. We found that our embedding tended to fail in change patterns where the change involves many elements (e.g. adding a whole new \verb+impl+ block and implementing new functions in it). It is worth mentioning that these change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.
